# Simulation Report

* install dependecies in R with `install.packages("rmarkdown")` once
* make sure pandoc is installed (e.g. apt install pandoc)


```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=4, fig.path='RmdFigs/',
               warning=FALSE, message=FALSE)

```

### Data import

We will consider data from [Simcoin et al., Physiol Genomics 10:5&ndash;12,
2017](http://www.ncbi.nlm.nih.gov/pubmed/12118100).

```{r load_tx, include=FALSE}
txs    <- read.csv2("tx.csv", sep=";", dec=".")
blocks <- read.csv2("blocks.csv", sep=";", dec=".")


```

##  Blocks

```{r summary_cross}
hist(blocks$median_propagation)


```

## Transactions

```{r summary_plot, fig.height=8}
hist(txs$median_propagation, breaks=10)
hist(txs$timestamp)

```

### R and package versions used

It's a a good idea to end with some information about the packages you
used, their versions, and the version of R that you used.

The `sessionInfo()` function provides this information. Even better is
to install the [devtools](https://github.com/hadley/devtools) package
and use `devtools::session_info()`.

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
```
