---
title: "Simcoin run Report"
date: '`r format(Sys.time(), "%d %B, %Y")`'
author: "simcoin"
header-includes:
   - \usepackage{float}
output:
  pdf_document:
    pandoc_args:
      - -V
      - classoption=twocolumn
    toc: yes
---

```{r knitr_options, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  fig.path='RmdFigs/',
  warning=FALSE,
  message=FALSE,
  error = TRUE,
  echo = FALSE,
  dev = 'pdf',
  fig.align='center'
  )

# set working directory for development: knitr::opts_knit$set(root.dir = '/tmp/run/postprocessing/')
```

```{r imports}
library(jsonlite)
library(anytime)
library(dplyr)
library(kableExtra)
library(lattice)
library(stringr)
Sys.setenv(TZ='UTC')
```

```{r preprocessing}
tick_infos <- read.csv("tick_infos.csv")
args       <- stream_in(file('../args.json'))

skip_ticks    <- ceiling(max(1/args$blocks_per_tick, 1/args$txs_per_tick, 1/args$tick_duration))
analyze_start <- tick_infos$start[skip_ticks + 1]
analyze_end   <- tick_infos$start[nrow(tick_infos) - skip_ticks - 1]

peer_logic_validation <- read.csv("peer_logic_validation_raw.csv")
update_tip            <- read.csv("update_tip_raw.csv")

blocks <- rbind(peer_logic_validation, update_tip[, !(names(update_tip) %in% c('tx', 'height'))])
blocks <- blocks %>% filter(timestamp > analyze_start & timestamp < analyze_end) %>% 
  group_by(hash) %>% filter(which.min(timestamp)==row_number())
write.csv(blocks, 'blocks.csv', row.names=FALSE, quote=FALSE)

files = c("blocks_create", "txs")
for (i in 1:length(files)) {
  data <- read.csv(paste(files[i], '_raw.csv', sep = ''))
  data <- data %>% filter(timestamp > analyze_start & timestamp < analyze_end)
  write.csv(data, paste(files[i], 'csv', sep = '.'), row.names=FALSE, quote=FALSE)
}

files = c("blocks_reconstructed", "txs_received", "blocks_received", "peer_logic_validation", "update_tip")
for (i in 1:length(files)) {
  data <- read.csv(paste(files[i], '_raw.csv', sep = ''))
  data <- data %>% filter(timestamp > analyze_start)
  write.csv(data, paste(files[i], 'csv', sep = '.'), row.names=FALSE, quote=FALSE)
}
```

```{r setup, include=FALSE}

blocks                <- read.csv("blocks.csv", dec=".")
blocks_create         <- read.csv("blocks_create.csv", dec=".")
blocks_reconstructed  <- read.csv("blocks_reconstructed.csv", dec=".")
blocks_received       <- read.csv("blocks_received.csv", dec=".")
block_exces           <- read.csv("block_exceptions.csv", dec=".")
peer_logic_validation <- read.csv("peer_logic_validation.csv", dec=".")
update_tip            <- read.csv("update_tip.csv", dec=".")
consensus_chain       <- read.csv("consensus_chain.csv", dec=".")

txs          <- read.csv("txs.csv", dec=".")
txs_received <- read.csv("txs_received.csv", dec=".")
tx_exces     <- read.csv("tx_exceptions.csv", dec=".")

rpc_exces <- read.csv("rpc_exceptions.csv", dec=".")
tips      <- read.csv("tips.csv", dec=".")

general_infos <- fromJSON(file('general_infos.json'))
step_times    <- read.csv("step_times.csv", dec=".")
tick_infos    <- read.csv("tick_infos.csv", dec=".")

memory   <- read.csv("memory.csv", dec=".")
cpu_time <- read.csv("cpu_time.csv", dec=".")

log_errors  <- readChar("log_errors.txt", file.info("log_errors.txt")$size)

ticks     <- readLines('../ticks.csv')
args      <- stream_in(file('../args.json'))
sim_start <- tick_infos$start[1]
```

## Simulation

```{r general_infos}
general_infos_table <- data.frame(
                    c(format(utctime(step_times$timestamp[1]), '%Y-%m-%d %H:%M:%S %Z'),
                      format(utctime(step_times$timestamp[length(step_times)]), '%Y-%m-%d %H:%M:%S %Z'),
                      general_infos$current_commit)
                  )
row.names(general_infos_table)  <- c('Start', 'End', 'Commit')
kable(general_infos_table,
      col.names = c('Value'), caption = 'General information', format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r start_arguments}
args_transposed <- as.data.frame(t(args))
kable(args_transposed, col.names = c('Value'), caption = 'Start arguments', format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r events_summary}
planned_blocks <- sum(str_count(ticks, 'block'))
planned_txs    <- sum(str_count(ticks, 'tx'))
df             <- data.frame(
                    c(length(ticks), nrow(tick_infos)),
                    c(planned_blocks, nrow(blocks_create)),
                    c(planned_txs, nrow(txs) - nrow(blocks_create))
                  )
row.names(df)  <- c('Planned', 'Run')
kable(df,
      col.names = c('Ticks', 'Blocks', 'Transactions'), caption = 'Events summary', format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

## System

```{r system_information}
kable(Sys.info(), col.names = c('Value'), format = 'latex', caption = 'System information') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r system_hardware_specs}
infos <- data.frame(c(general_infos$cpu_model, general_infos$cpus, round(as.double(general_infos$total_memory) / 1000000, 3)))
row.names(infos) <- c('CPU model', 'CPU(s)', 'Memory [GB]')
kable(infos, col.names = c('Value'), format = 'latex', caption = 'System hardware specs') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r cpu_usage_over_time, fig.cap='CPU usage over time', fig.env='figure*', fig.height=3}
cpu_time$total          <- cpu_time$user + cpu_time$nice + cpu_time$system
cpu_time_diff           <- data.frame(cpu_time$timestamp[-1], diff(cpu_time$total), diff(cpu_time$idle))
colnames(cpu_time_diff) <- c('timestamp', 'total', 'idle')
cpu_time_diff$usage     <- (1 - cpu_time_diff$idle / (cpu_time_diff$total + cpu_time_diff$idle)) * 100

xyplot(cpu_time_diff$usage ~ cpu_time_diff$timestamp - sim_start, xlab = 'Time [s]', ylab = 'CPU usage [%]', t = 'l')
```

```{r memory_usage_over_time, fig.cap='Memory usage over time', fig.env='figure*', fig.height=3}
memory$usage <- (1 - memory$available/memory$total) * 100
xyplot(memory$usage ~ memory$timestamp - sim_start, xlab = 'Time [s]', ylab = 'Memory usage [%]', t = 'l' )
```

## Ticks

```{r simulation}
step_diff                      <- data.frame(diff(step_times$timestamp))
colnames(step_diff)            <- c('timestamp')
step_diff_summary              <- rbind(step_diff, apply(step_diff, 2, sum))
step_diff_summary$duration_min <- step_diff_summary$timestamp/60
step_diff_summary              <- step_diff_summary %>%
  mutate_at(.vars = vars(timestamp, duration_min), .funs = funs(round(.,2)))
rownames(step_diff_summary)    <- c('Preparation step', 'Simulation step', 'Postprocessing step', 'Total')

kable(step_diff_summary, col.names = c('Duration [s]', 'Duration [min]'), caption = 'Duration of different steps', format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r step_times_barchart, fig.cap='Step times', fig.pos='H', fig.height=3}
step_diff             <- data.frame(diff(step_times$timestamp))
t_step_diff           <- t(step_diff)
colnames(t_step_diff) <- c('Preparation', 'Simulation', 'Postprocessing')
rownames(t_step_diff) <- c('Run')
barchart(t_step_diff, stack = TRUE, ref = FALSE, xlab = 'Time [s]',
         auto.key = list(columns=2, rectangles = TRUE, points = FALSE, space='top', text=colnames(t_step_diff)),
         panel=function(x,y,...){
      panel.barchart(x,y,...)
      xx <- unsplit(unname(lapply(split(x, y), function(t)cumsum(t)-t/2)), y)
      ltext(xx, y=y, labels=round(x, 1))
})
```

```{r tick_duration_summary}
kable(unclass(summary(tick_infos$duration)), col.names = c('Duration [s]'), digits = 3, caption = 'Tick duration', format = 'latex')  %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r tick_duration_boxplot, fig.cap='Tick duration', fig.pos='H', fig.height=3}
bwplot(tick_infos$duration, xlab='Tick duration [s]')
```

```{r tick_duration_xyplot, fig.cap='Tick duration over time', fig.env='figure*'}
xyplot(tick_infos$duration ~ tick_infos$start - sim_start, xlab = 'Time [s]', ylab = 'Duration [s]', t = 'l' )
```

```{r tick_duration_median, fig.cap='Tick duration over time with boxplot', fig.env='figure*'}
# par(mar=c(7,5,1,1)) # increase plot margins to fit longer labels on the x-axis
# boxplot(tick_infos$duration ~ cut(tick_infos$start - sim_start, 25), las=2)
```

##  Blocks

```{r blocks_setup}
block_propagation <- rbind(blocks_reconstructed, blocks_received, peer_logic_validation, update_tip[, !(names(update_tip) %in% c('tx', 'height'))])

block_propagation <- merge(block_propagation, blocks$hash)[colnames(block_propagation)]

block_propagation <- block_propagation %>%
  group_by(node, hash) %>% filter(which.min(timestamp)==row_number()) %>%
  group_by(hash) %>% mutate(propagation_time = timestamp - min(timestamp)) %>%
  filter(propagation_time != 0)
```

`r if(nrow(blocks_create) == 0) {'There were 0 blocks during this simulation run.'}`

```{r block_stale_barchart, fig.cap='Accepted/stale blocks', fig.pos='H', fig.height=3}
consensus_chain$stale = 'Accepted'
blocks <- merge(blocks, consensus_chain, all.x = TRUE)
blocks$stale[is.na(blocks$stale)] <- 'Stale'

stale_block <- t(table(blocks$stale))
rownames(stale_block) <- c('Run')
barchart(stale_block, stack = TRUE, ref = FALSE, xlab = 'Blocks',
         auto.key = list(columns=2, points = FALSE, rectangles = TRUE), panel=function(x,y,...){
       panel.barchart(x,y,...)
       xx <- do.call('c', unname(lapply(split(x, y), function(t)cumsum(t)-t/2)))
       ltext(xx, y=y, labels=x)
})
```

```{r block_stale_rate_summary}
factor_stale_block <- t(table(factor(blocks$stale, levels = c('Accepted', 'Stale'))))
stale_rate <- data.frame(c(round(prop.table(factor_stale_block)[,2] * 100, 3)))
rownames(stale_rate) <- c('Block stale rate')
kable(stale_rate, col.names = c('Rate [%]'), digits = 3, caption = 'Stale block rate', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r block_size_summary}
kable(unclass(summary(blocks_create$total_size / 1000)), col.names = c('Size [kB]'), digits = 3, caption = 'Block size', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r block_size_boxplot, fig.cap='Block size', fig.pos='H', fig.height=3}
bwplot(blocks_create$total_size / 1000, xlab='Block size [kb]')
```

```{r block_propagation_summary}
kable(unclass(summary(block_propagation$propagation_time * 1000)), col.names = c('Propagation [ms]'), digits = 3, caption = 'Block propagation', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r block_propagation_density, fig.cap='Block propagation density', fig.env='figure*'}
densityplot(~propagation_time, data=block_propagation, auto.key = list(space = "right"), ref = TRUE, xlab = 'Duration [s]', plot.points = FALSE)
```

## Transactions

`r if(nrow(txs) == 0) {'There were 0 transactions during this simulation run.'}`

```{r txs_propagation_summary}
txs_received <- merge(txs[, c('timestamp', 'hash')], txs_received, by = 'hash')
txs_received$propagation_duration <- txs_received$timestamp.y - txs_received$timestamp.x

kable(unclass(summary(txs_received$propagation_duration * 1000)), col.names = c('Propagation [ms]'), digits = 3, caption = 'Transactions propagation', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r tx_propagation_density, fig.cap='Transaction propagation density', fig.env='figure*'}
densityplot(~propagation_duration, data=txs_received, auto.key = list(space = "right"), ref = TRUE, xlab = 'Duration [s]', plot.points = FALSE)
```

```{r txs_per_block_summary}
kable(unclass(summary(blocks_create$txs)), digits = 2, caption = 'Transactions per block', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r txs_per_block_boxplot, fig.cap='Transactions in block', fig.pos='H', fig.height=3}
bwplot(blocks_create$txs, xlab='Transactions')
```

## Tips

```{r tips_setup}
tips          <- tips[tips$status != 'active',]
tips_per_node <- data.frame(as.matrix(table(tips$node)))
colnames(tips_per_node) <- c('count')
```

`r if(nrow(tips) == 0) {'There were 0 tips during this simulation run.'}`

```{r tips_per_node_summary, eval=nrow(tips) > 0}
kable(unclass(summary(tips_per_node$count)), digits = 2, caption = 'Tips per node', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r tips_per_node_boxplot, eval=nrow(tips) > 0, fig.cap='Tips per node', fig.pos='H', fig.height=3}
bwplot(tips_per_node$count, xlab='Tips')
```

```{r tips_branchlen_summary, eval=nrow(tips) > 0}
kable(unclass(summary(tips$branchlen)), digits = 2, caption = 'Overall branch length of tips', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r tips_branchlen_boxplot, eval=nrow(tips) > 0, fig.cap='Overall branch length of tips', fig.pos='H', fig.height=3}
bwplot(tips$branchlen, xlab='Branch length')
```

## Exceptions
```{r event_exces_setup}
if (nrow(tx_exces) > 0) {
  tx_exces$type    <- 'Transaction'
}
if (nrow(block_exces) > 0) {
  block_exces$type <- 'Block'
}
event_exces <- rbind(tx_exces, block_exces)
```

`r if(nrow(event_exces) == 0) {'There were 0 exceptions during the creation of transaction/block events.'}`

```{r event_exces, eval=nrow(event_exces) > 0, fig.cap='Exceptions transaction/block event', fig.pos='H', fig.height=3}
barchart(table(factor(event_exces$type, levels = c('Transaction', 'Block'))), ref = FALSE, xlab = 'Amount of exceptions',
         panel=function(x,y,...){
          panel.barchart(x,y,...)
          ltext(x=x, y=y, labels=x, pos = 4)
})
```

```{r tx_exces_type, eval=nrow(tx_exces) > 0, fig.cap='Transaction exception types', fig.pos='H', fig.height=nrow(tx_exces)}
barchart(table(tx_exces$exception), ref = FALSE, xlab = 'Amount of exceptions',
         panel=function(x,y,...){
          panel.barchart(x,y,...)
          ltext(x=x, y=y, labels=x, pos = 4)
})
```

```{r block_exces_type, eval=nrow(block_exces) > 0, fig.cap='Block exception types', fig.pos='H', fig.height=nrow(block_exces)}
barchart(table(block_exces$exception), ref = FALSE, xlab = 'Amount of exceptions',
         panel=function(x,y,...){
          panel.barchart(x,y,...)
          ltext(x=x, y=y, labels=x, pos = 4)
})
```

```{r rpc_exces, eval=nrow(rpc_exces) > 0, fig.cap='RPC exceptions', fig.pos='H', fig.height=length(unique(rpc_exces$method)) + 2}
rpc_exces_table <- table(rpc_exces$method, rpc_exces$exception)
barchart(rpc_exces_table, ref = FALSE, xlab = 'Amount of exceptions',
         auto.key = list(columns=2, rectangles = TRUE, points = FALSE, space='top', text=colnames(rpc_exces_table)),
         panel=function(x,y,...){
           panel.barchart(x,y,...)
              xx <- unsplit(unname(lapply(split(x, y), function(t)cumsum(t)-t/2)), y)
              xx <-  xx [! x %in% 0]
              y <-  y [! x %in% 0]
              x <-  x [! x %in% 0]
              ltext(xx, y=y, labels=x)
})
```
\clearpage

#### R and package versions used

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
