# Simulation Report

* install dependecies in R with `install.packages("dependency")`
    * rmarkdown
    * devtools
* make sure pandoc is installed (e.g. apt install pandoc)


```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=4, fig.path='RmdFigs/',
               warning=FALSE, message=FALSE)
```

### Data import

We will consider data from [Simcoin et al., Physiol Genomics 10:5&ndash;12,
2017](http://www.ncbi.nlm.nih.gov/pubmed/12118100).

```{r setup, include=FALSE}
txs    <- read.csv2("tx.csv", sep=";", dec=".")
blocks <- read.csv2("blocks.csv", sep=";", dec=".")
txs_count <- nrow(txs)
blocks_count <- nrow(blocks)
```

##  Blocks

There were `r blocks_count` blocks in this simulation.

```{r blocks, eval=blocks_count > 0}
hist(blocks$median_propagation)
```

## Transactions

There were `r txs_count` txs in this simulation.

```{r txs, eval=txs_count > 0, fig.height=8}
hist(txs$median_propagation, breaks=10)
hist(txs$timestamp)
```


### R and package versions used

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
