---
title: "Simcoin run-report"
header-includes:
   - \usepackage{float}
output:
  pdf_document:
    pandoc_args: [
      "-V", "classoption=twocolumn"
    ]
    toc: yes
date: '`r format(Sys.time(), "%d %B, %Y")`'
author: "simcoin"
---

```{r setup, include=FALSE}
library(knitr)
library(jsonlite)
library(anytime)
library(dplyr)
library(kableExtra)
Sys.setenv(TZ='UTC')

knitr::opts_chunk$set(
  fig.path='RmdFigs/',
  warning=FALSE,
  message=FALSE,
  error = TRUE,
  echo = FALSE,
  dev = 'pdf',
  fig.align='center'
  )

# set working directory for development: knitr::opts_knit$set(root.dir = '/tmp/run/postprocessing/')

# general infos is needed twice to avoid exception in simulation section.
general_infos <- fromJSON(file('general_infos.json'))
general_info2 <- fromJSON(file('general_infos.json'))
tick_infos    <- read.csv2("tick_infos.csv", dec=".")

txs    <- read.csv2("txs.csv", dec=".")
blocks <- read.csv2("blocks.csv", dec=".")
tips   <- read.csv2("tips.csv", dec=".")

memory   <- read.csv2("memory.csv", dec=".")
cpu_time <- read.csv2("cpu_time.csv", dec=".")

blocks_received <- read.csv2("blocks_received.csv", dec=".")
txs_received    <- read.csv2("txs_received.csv", dec=".")

tx_exces    <- read.csv2("tx_exceptions.csv", dec=".")
block_exces <- read.csv2("block_exceptions.csv", dec=".")
rpc_exces   <- read.csv2("rpc_exceptions.csv", dec=".")
log_errors  <- readChar("log_errors.txt", file.info("log_errors.txt")$size)

args      <- stream_in(file('../args.json'))
sim_start <- tick_infos$start[1]
```

## Simulation

The simulation was started with the start arguments shown by table \ref{tab:start_arguments} at
**`r format(utctime(general_info2$step_times[['preparation_start']]), '%Y-%m-%d %H:%M:%S %Z')`**
and ended at
**`r format(utctime(general_info2$step_times[['postprocessing_end']]), '%Y-%m-%d %H:%M:%S %Z')`**.
During the simulation `r if(nrow(tick_infos) == args['amount_of_ticks']) {'all'} else {'only'}` **`r nrow(tick_infos)` of `r args['amount_of_ticks']`** ticks were executed in time, where one tick took `r args['tick_duration']` second(s).
The checked out commit at execution time was:

`r general_infos$current_commit`

```{r start_arguments}
args_transposed <- as.data.frame(t(args))
kable(args_transposed, col.names = c('value'), caption = 'Start arguments', format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r simulation}
step_times           <- data.frame(t(data.frame(general_infos$step_times)))
colnames(step_times) <- c('timestamp')
step_times           <- step_times[order(step_times$timestamp), , drop = FALSE]

diff_steps <- data.frame(diff(as.matrix(step_times)))
diff_steps <- rbind(diff_steps, apply(diff_steps, 2, sum))
diff_steps$duration_min <- diff_steps$timestamp/60
diff_steps <- diff_steps %>% mutate_at(.vars = vars(timestamp, duration_min), .funs = funs(round(.,2)))
rownames(diff_steps)   <- c('preparation step', 'simulation step', 'postprocessing step', 'total')

kable(diff_steps, col.names = c('duration [s]', 'duration [min]'), caption = 'Duration of different steps', format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r ticks}
tick_infos$rel_timestamp <- tick_infos$start - sim_start
```

```{r tick_duration_summary}
kable(unclass(summary(tick_infos$duration)), col.names = c('duration [s]'), digits = 3, caption = 'Tick duration', format = 'latex')  %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r tick_duration_boxplot, fig.cap='Tick duration', fig.pos='H'}
boxplot(tick_infos$duration, ylab='Tick duration [s]')
```

```{r tick_duration_median, fig.cap='Tick duration over time', fig.env='figure*'}
par(mar=c(7,5,1,1)) # increase plot margins to fit longer labels on the x-axis
boxplot(tick_infos$duration ~ cut(tick_infos$rel_timestamp, 25), las=2)
```

## System
```{r base_information}
infos <- data.frame(c(general_infos$hostname, general_infos$cpu_model, general_infos$cpus, round(as.double(general_infos$total_memory) / 1000000, 3)))
row.names(infos) <- c('hostname', 'CPU model', 'CPU(s)', 'memory [GB]')
kable(infos, col.names = c('value'), format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
```

```{r system}
memory$rel_timestamp <- memory$timestamp - sim_start
memory$usage         <- (1 - memory$free/memory$total) * 100

cpu_time$rel_timestamp   <- cpu_time$timestamp - sim_start
cpu_time$total           <- cpu_time$user + cpu_time$nice + cpu_time$system
cpu_time_delta           <- data.frame(cpu_time$rel_timestamp[-1], diff(cpu_time$total), diff(cpu_time$idle))
colnames(cpu_time_delta) <- c('rel_timestamp', 'total', 'idle')
cpu_time_delta$usage     <- (cpu_time_delta$idle / cpu_time_delta$total) * 100
```

During the run every `r args['system_snapshots_frequency']` seconds the files `/proc/stat` and `/proc/meminfo` were read to monitor the system. In total 100 snapshots were collected. Figure A shows the CPU utilization and figure B shows the memory usage during the run.

```{r memory_usage_over_time, fig.cap='Memory usage over time', fig.env='figure*'}
plot(memory$rel_timestamp, memory$usage, xlab = 'Time [s]', ylab='Memory usage [%]', type = 'b')
```

```{r cpu_usage_over_time, fig.cap='CPU usage over time', fig.env='figure*'}
plot(cpu_time_delta$rel_timestamp, cpu_time_delta$usage, xlab = 'Time [s]', ylab='CPU usage [%]', type = 'b')
```

##  Blocks
In this simulation `r nrow(blocks)` blocks were created by `r length(unique(blocks$node))` different nodes. Note: Some blocks may never reached some nodes because of forks in the chain or other reasons.


```{r blocks}
blocks$rel_timestamp <- blocks$timestamp - sim_start
```

```{r blocks_distribution, fig.cap='Block creators', fig.pos='H'}
blocks_per_node <- table(blocks$node)
percentage <- round(blocks_per_node/sum(blocks_per_node)*100)
labels <- paste(names(blocks_per_node), ' blocks: ', blocks_per_node, ', ', percentage, '%', sep='')
pie(blocks_per_node, labels)
```

```{r stale, fig.cap='Accepted/stale blocks', fig.pos='H'}
stale_blocks <- table(blocks$stale)
percentage <- round(stale_blocks/sum(stale_blocks)*100)
labels <- paste(names(stale_blocks), ' blocks: ', stale_blocks, ', ', percentage, '%', sep='')
pie(stale_blocks, labels)
```

```{r block_size_summary}
kable(unclass(summary(blocks$total_size / 1000)), col.names = c('size [kB]'), digits = 3, caption = 'Block size', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r block_size_boxplot, fig.cap='Block size', fig.pos='H'}
boxplot(blocks$total_size / 1000, ylab='Block size [kb]')
```

```{r block_propagation_summary}
kable(unclass(summary(blocks_received$propagation_duration * 1000)), col.names = c('propagation [ms]'), digits = 3, caption = 'Block propagation', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r block_propagation_density, fig.cap='Block propagation density', fig.env='figure*'}
hist(blocks_received$propagation_duration, xlab = 'Time [s]', main = NULL)
lines(density(blocks_received$propagation_duration))
```

## Transactions
There were `r nrow(txs)` transaction created in this simulation. Note: Some transactions may reached other nodes included in a block. Hence they wouldn't appear in this statistics.


```{r txs}
txs$rel_timestamp <- txs$timestamp - sim_start
```

```{r txs_distribution, fig.cap='Transaction creators', fig.pos='H'}
txs_per_node <- table(txs$node)
percentage <- round(txs_per_node/sum(txs_per_node)*100)
labels <- paste(names(txs_per_node), ' blocks: ', txs_per_node, ', ', percentage, '%', sep='')
pie(txs_per_node, labels)
```

```{r txs_propagation_summary}
kable(unclass(summary(txs_received$propagation_duration * 1000)), col.names = c('propagation [ms]'), digits = 3, caption = 'Transactions propagation', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r tx_propagation_density, fig.cap='Transaction propagation density', fig.env='figure*'}
hist(txs_received$propagation_duration, xlab = 'Time [s]', main = NULL)
lines(density(txs_received$propagation_duration))
```

```{r txs_per_block_summary}
kable(unclass(summary(blocks$txs)), digits = 0, caption = 'Transactions per block', format = 'latex') %>%
  kable_styling(latex_options = c('HOLD_position'))
```

```{r txs_per_block_boxplot, fig.cap='Transactions in block', fig.pos='H'}
boxplot(blocks$txs, ylab='Transactions')
```

## Nodes
In this simulation were `r length(unique(tips$node))` nodes. In the following the statistics based on data from the RPC-call `getchaintips` are presented.

```{r tips}
kable(tips, format = 'latex') %>%
  kable_styling(latex_options = c('scale_down','HOLD_position'))
not_active_tips <- tips[!(tips$status=='active'),]
```

## Block exceptions
There were `r nrow(block_exces)` exceptions catched during the execution of the RPC-call to create a block.

```{r block_exces_hist, eval=nrow(block_exces), fig.pos='H'}
hist(block_exces$rel_timestamp - sim_start, xlab = 'Time [s]')
```

```{r block_exces_node_distribution, eval=nrow(block_exces), fig.pos='H'}
exces_per_node <- table(block_exces$node)
percentage     <- round(exces_per_node/sum(exces_per_node)*100)
labels         <- paste(names(exces_per_node), ': ', exces_per_node, ', ', percentage, '%', sep='')
pie(exces_per_node, labels)
title(main = 'Distribution of exceptions')
```

```{r block_exces_distribution, eval=nrow(block_exces), , fig.pos='H'}
exces_per_node <- table(block_exces$exception)
percentage     <- round(exces_per_node/sum(exces_per_node)*100)
labels         <- paste(names(exces_per_node), ': ', exces_per_node, ', ', percentage, '%', sep='')
pie(exces_per_node, labels)
title(main = 'Exception types')
```

## Transaction exceptions
There were `r nrow(tx_exces)` exceptions catched during the execution of the RPC-call to send a transaction.

```{r tx_exces, eval=nrow(tx_exces), fig.pos='H'}
tx_exces$rel_timestamp <- tx_exces$timestamp - sim_start
hist(tx_exces$rel_timestamp, xlab = 'Time [s]')
```

```{r tx_exces_node_distribution, eval=nrow(tx_exces), fig.pos='H'}
exces_per_node <- table(tx_exces$node)
percentage     <- round(exces_per_node/sum(exces_per_node)*100)
labels         <- paste(names(exces_per_node), ': ', exces_per_node, ', ', percentage, '%', sep='')
pie(exces_per_node, labels)
title(main = 'Distribution of exceptions')
```

```{r tx_exces_distribution, eval=nrow(tx_exces), fig.pos='H'}
exces_per_node <- table(tx_exces$exception)
percentage     <- round(exces_per_node/sum(exces_per_node)*100)
labels         <- paste(names(exces_per_node), ': ', exces_per_node, ', ', percentage, '%', sep='')
pie(exces_per_node, labels)
title(main = 'Exception types')
```

## RPC exceptions
There were `r nrow(rpc_exces)` exceptions catched during a RPC-call.

```{r rpc_exces, eval=nrow(rpc_exces), fig.pos='H'}
rpc_exces$rel_timestamp <- rpc_exces$timestamp - sim_start
hist(rpc_exces$rel_timestamp, xlab = 'Time [s]')
```

```{r rpc_exces_node_distribution, eval=nrow(rpc_exces), fig.pos='H'}
exces_per_node <- table(rpc_exces$node)
percentage     <- round(exces_per_node/sum(exces_per_node)*100)
labels         <- paste(names(exces_per_node), ': ', exces_per_node, ', ', percentage, '%', sep='')
pie(exces_per_node, labels)
title(main = 'Distribution of exceptions')
```

```{r rpc_exces_distribution, eval=nrow(rpc_exces), eval=nrow(rpc_exces), fig.pos='H'}
exces_per_node <- table(rpc_exces$exception)
percentage     <- round(exces_per_node/sum(exces_per_node)*100)
labels         <- paste(names(exces_per_node), ': ', exces_per_node, ', ', percentage, '%', sep='')
pie(exces_per_node, labels)
title(main = 'Exception type')
```

\clearpage

#### R and package versions used

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
